Data Specialization for a Staged Lambda Calculus

### 2015-02-26 paper/paper.tex
minor edits
dangerous and experimental new masking section

\lstset{%language=ML,
        columns=fullflexible,
        keepspaces=true,
        basicstyle=\ttfamily,
        tabsize=4,
%        escapeinside={"*}{*"},
        escapeinside={@}{@},
        moredelim=*[is][\color{stageOneColor}]{1`}{`},
        moredelim=*[is][\color{stageTwoColor}]{2`}{`},
        moredelim=*[is][\color{stageMonoColor}]{3`}{`},
        literate={|->}{{\ $\mapsto$\ }}1
                 {xhat}{{$\hat{\texttt{x}}$}}1
                 {yhat}{{$\hat{\texttt{y}}$}}1
                 {zhat}{{$\hat{\texttt{z}}$}}1
                 {murec}{{$\mu$}}1
                 {alpha}{{$\alpha$}}1
                 {beta}{{$\beta$}}1
                 {atsign}{{@}}1
                 {literalone}{1}1 % hax
                 {literaltwo}{2}1 % hax
                 {^}{{$\curr$}}1
                 {$}{{$\fut$}}1}

>>>>

\lstset{%language=ML,
        columns=fullflexible,
        keepspaces=true,
        basicstyle=\ttfamily,
        tabsize=4,
%        escapeinside={"*}{*"},
        escapeinside={@}{@},
        moredelim=*[is][\color{stageOneColor}]{1`}{`},
        moredelim=*[is][\color{stageTwoColor}]{2`}{`},
        moredelim=*[is][\color{stageMonoColor}]{3`}{`},
        moredelim=*[is][\colorbox{black}]{4`}{`},
        literate={|->}{{\ $\mapsto$\ }}1
                 {xhat}{{$\hat{\texttt{x}}$}}1
                 {yhat}{{$\hat{\texttt{y}}$}}1
                 {zhat}{{$\hat{\texttt{z}}$}}1
                 {murec}{{$\mu$}}1
                 {alpha}{{$\alpha$}}1
                 {beta}{{$\beta$}}1
                 {atsign}{{@}}1
                 {literalone}{1}1 % hax
                 {literaltwo}{2}1 % hax
                 {^}{{$\curr$}}1
                 {$}{{$\fut$}}1}

### 2015-02-26 paper/splitting.tex
dangerous and experimental new masking section

The rules of the masking relation are given in \ref{fig:valMask}.
Masking operates by first inducting on the entries of the residual table.  
Being purely second-stage content, these are reified into let statements at the top of the resumer.
Once the table is empty, masking inducts on value itself.

Masking assigns ground values to the immediate value
and likewise assigns references into the residual table to the resumer.
In both cases, the alternate component is assigned to $\tup{}$, to represent trivial information.
Note that the \texttt{grnd} and \texttt{next} annotations are erased.

Masking distributes into tuples, injections, and rolls, since their subvalues may have content at both stages.
However, the tags of injections and rolls are replicated only in the immediate value, 
since they represent first-stage information.

Since lambdas may represent multistage computations, 
masking splits the body of lambdas as general world \bbonem\ terms (as described in \ref{sec:split-one}), 
and packages the resulting terms as functions.

>>>>

So to motivate the definition we chose, consider the following residual table and partial value,
colored as before:
\begin{lstlisting}
2`[yhat |-> 1+2]`1` (next{`2`yhat`1`}, injL (grnd{7}, next{`2`yhat`1`}))`
\end{lstlisting}
To reduce this term to just its first-stage information ($i$ in the output), we start by redacting all of the second stage (blue) parts,
along with the surrounding \texttt{next} annotations:
\begin{lstlisting}
4`##########`1` (`4`######`1`, injL (grnd{7}, `4`######`1`))`
\end{lstlisting}
The holes inside the term needs to be filled with something to convey the lack of information, so we use a unit values:
\begin{lstlisting}
1`((), injL (grnd{7},()))`
\end{lstlisting}
Finally, we drop \texttt{grnd} annotations:
\begin{lstlisting}
1`((), injL (7,()))`
\end{lstlisting}
To reduce the original partial value to just its second-stage information ($q$ in the output), 
we start by redacting all \texttt{grnd} blocks, \texttt{next} annotations, and injection tags
(essentially, the tuple structure is kept, but all other structure is discarded):
\begin{lstlisting}
2`[yhat |-> 1+2]`1` (`4`####`2`yhat`4`#`1`, `4`####`1` (`4`####`1`,`4`####`2`yhat`4`#`1`))`
\end{lstlisting}
Where the \texttt{grnd} blocks left a hole, we put in a unit values:
\begin{lstlisting}
2`[yhat |-> 1+2] (yhat, ((),yhat))`
\end{lstlisting}
Finally, we reify the residual table into \texttt{let} bindings:
\begin{lstlisting}
2`let yhat = 1+2 in (yhat, ((),yhat))`
\end{lstlisting}

The formal implementation of the masking relation is given in \ref{fig:valMask}.
Note that since lambdas may represent multistage computations, 
masking splits the body of lambdas as general world \bbonem\ terms (as described in \ref{sec:split-one}), 
and packages the resulting terms as functions in both the first and second stage.
% Masking operates by first inducting on the entries of the residual table.  
% Being purely second-stage content, these are reified into let statements at the top of the resumer.
% Once the table is empty, masking inducts on value itself.

% Masking assigns ground values to the immediate value
% and likewise assigns references into the residual table to the resumer.
% In both cases, the alternate component is assigned to $\tup{}$, to represent trivial information.
% Note that the \texttt{grnd} and \texttt{next} annotations are erased.

% Masking distributes into tuples, injections, and rolls, since their subvalues may have content at both stages.
% However, the tags of injections and rolls are replicated only in the immediate value, 
% since they represent first-stage information.
